Regular singular Volterra equations on complex domains

### 2023-09-01 reg-sing-volterra.tex
no changes

We'll reduce Theorem~\ref{thm:general_volterra} to Lemma~\ref{lem:perturbed_volterra} in Section~\ref{sec:existence and uniqueness}, by rewriting the homogeneous equation we want to solve as an inhomogeneous equation in a more regular space. To show that the inhomogeneous term, $\softpart f_0$, is regular enough, we'll use Proposition~\ref{prop:smoothing} from Section~\ref{sec:image under soft_part} proving that $\softpart$ improves on the regularity of $f_0$ described by Theorem~\ref{thm:proto-growth}.

\begin{rmk}
When $\hardpart$ also satisfies Condition~\eqref{cond:reg-p}, Theorem~\ref{thm:general_volterra} can be restated to give a unique solution in the affine subspace
\[ \zeta^{\tau-1} + \singexpalg{\rho-1}(\domain) \]
when $\rho > \tau$ is low enough, as described in Proposition~\ref{prop:alt-general_volterra} from Section~\ref{sec:existence and uniqueness}.
\end{rmk}
\subsection{Acknowledgements}
We are grateful for the conversations with Kihyun Kim, Angeliki Menegaki, and Sung-Jin Oh that informed the early stages of this work. This paper is a result of the ERC-SyG project, Recursive and Exact New Quantum Theory (ReNewQuantum) which received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme under grant agreement No 810573. It was also made possible by the hospitality of the Institut des Hautes \'{E}tudes Scientifiques during a return visit by AF, funded by the Fondation Math\'{e}matique Jacques Hadamard under the \textit{Junior Scientific Visibility} program.
\section{Function spaces for holomorphic Volterra operators}\label{fn-spaces}
\subsection{Weighted holomorphic $L^{\infty}$ spaces}\label{sec:fn-space-defs}
Throughout this paper, as described in Section~\ref{motivation}, the ``position'' variable $\zeta$ will be the standard coordinate on $\C$. Take a simply connected open set $\Omega \subset \C$ that touches but doesn't contain $\zeta = 0$. Let $\cont(\Omega)$ be the space of continuous complex-valued functions on $\Omega$. Give $\cont(\Omega)$ the compact-open topology, recalling that this is the coarsest topology in which the seminorm $f \mapsto \sup_K |f|$ is continuous for every compact subset $K \subset \Omega$~\cite[Example~2.6 and \S 4 notes]{fnl-cpx-anal}. The holomorphic functions form a closed subspace $\holo(\Omega) \subset \cont(\Omega)$~\cite[Proposition~3.14]{fnl-cpx-anal}.

Fixing a real constant $\Lambda$, let's restrict our attention to holomorphic functions on $\Omega$ which are bounded by constant multiples of $e^{\Lambda|\zeta|}$. One might describe these functions as being uniformly of exponential type $\Lambda$.\footnote{Recall that a function is of exponential type $\Lambda$ if for every $\varepsilon>0$ there is a constant $A_\varepsilon$ (which depends on $\varepsilon$) such that $|f(x)|\leq A_\varepsilon e^{(\Lambda+\varepsilon)|x|}$. We instead require that there exists a uniform constant $A$ such that for every $\varepsilon>0$ the bound holds.} They form a space $\singexp{0}{\Lambda}(\Omega)$, which we'll equip with the norm $\|f\|_{0,\Lambda} = \sup_\Omega e^{-\Lambda|\zeta|}\,|f|$. With respect to the seminorm on $\holo(\Omega)$ given by a compact set $K \subset \Omega$, the inclusion map $\singexp{0}{\Lambda}(\Omega) \hookrightarrow \holo(\Omega)$ has norm $\sup_K e^{\Lambda |\zeta|}$. That means the inclusion is continuous.
\begin{prop}\label{exp-complete}
The space $\singexp{0}{\Lambda}(\Omega)$ is complete.
\end{prop}
\begin{proof}
Take a Cauchy sequence $f_1, f_2, f_3, \ldots \in \singexp{0}{\Lambda}(\Omega)$. The inclusion map $\singexp{0}{\Lambda}(\Omega) \hookrightarrow \holo(\Omega)$ is bounded with respect to each of the seminorms on $\holo(\Omega)$ given by $|f| \mapsto \sup_K |f|$ for compact subsets $K \subset \Omega$, so our sequence is Cauchy in $\holo(\Omega)$ too. Since $\holo(\Omega)$ is complete~\cite[Proposition~3.5]{fnl-cpx-anal},\footnote{That is, a sequence which is Cauchy in each of the seminorms on $\holo(\Omega)$ will always converge in the topology of $\holo(\Omega)$, which is the coarsest topology in which all of the seminorms are continuous.} our sequence converges to a function $f$ there.

The Cauchy property in $\singexp{0}{\Lambda}(\Omega)$ tells us that for any $r > 0$, we can find some $n$ for which $e^{-\Lambda |\zeta|}\,|f_k - f_n| \le r$ whenever $k \ge n$. Since convergence in $\holo(\Omega)$ implies pointwise convergence, we can see as $k$ grows that $e^{-\Lambda |\zeta|}\,|f - f_n| \le r$. This shows that our sequence converges to $f$ in the norm $\|\cdot\|_{0,\Lambda}$. We can also see from this argument that $f$ is in $\singexp{0}{\Lambda}$: picking some $r > 0$, we observe that
\begin{align*}
e^{-\Lambda |\zeta|}\,|f| & \le e^{-\Lambda |\zeta|}\,|f - f_n| + e^{-\Lambda |\zeta|}\,|f_n| \\
& \le r + \|f_n\|_{0,\Lambda}
\end{align*}
for the corresponding $n$, showing that $e^{-\Lambda |\zeta|}\,|f|$ is bounded.
\end{proof}

Now, let's relax our norm to allow both exponential growth at infinity and a power-law singularity at $\zeta = 0$. Let $\singexp{\sigma}{\Lambda}(\Omega)$ be the space of holomorphic functions on $\Omega$ which are bounded by constant multiples of $|\zeta|^\sigma e^{\Lambda|\zeta|}$. Give it the norm $\|f\|_{\sigma,\Lambda} = \sup_\Omega |\zeta|^{-\sigma} e^{-\Lambda|\zeta|}\,|f|$. Reprising the arguments from above, we can show that the inclusion $\singexp{\sigma}{\Lambda}(\Omega) \hookrightarrow \holo(\Omega)$ is continuous, and we can generalize Proposition~\ref{exp-complete}:

@@

As we mentioned in Section~\ref{sec:fn-space-defs}, one might describe $\singexp{0}{\Lambda}(\domain)$ as the space of functions on $\domain$ which are uniformly of exponential type $\Lambda$. Taking the union of these spaces over all $\Lambda \in \R$, we get the space $\singexpalg{0}(\domain)$ that contains all functions of exponential type. Having continuous inclusions between the subspaces $\singexp{0}{\Lambda}(\domain)$ as $\Lambda$ increases, we can give $\singexpalg{0}(\domain)$ a meaningful topology: the finest topology that makes all the inclusions $\singexp{0}{\Lambda}(\domain) \hookrightarrow \singexpalg{0}(\domain)$ continuous.\footnote{In category-theoretic language, $\singexpalg{0}(\domain)$ is the limit of the family $\singexp{0}{\Lambda}(\domain)$.}

@@

When $\sigma' > \sigma$, there's a continuous inclusion $\singexpalg{\sigma'}(\Omega)\hookrightarrow \singexpalg{\sigma}(\Omega)$.

@@

\begin{prop}\label{prop:get-contraction}
Suppose that $\hardpart$ satisfies {\em Conditions \eqref{cond:sing}} and \eqref{cond:diag-basic}, and $\softpart$ satisfies {\em Condition \eqref{cond:eps-lambda}}. Then, for each $\rho > \tau$, we can ensure that $\volterra$ is a contraction of $\singexp{\rho-1}{\Lambda}$ by making $\Lambda$ big enough.
\end{prop}
\begin{rmk}
For the argument, we'll use, ``big enough'' always requires $\Lambda > \lambda_\Delta$, and may require $\Lambda \gg \lambda_\Delta$.
\end{rmk}
First, pick some $\sigma \in (\tau, \rho)$.\footnote{If you'd like, you can fix a choice of $\sigma$---for example, taking $\sigma = \frac{1}{2}(\tau + \rho)$.} By Proposition~\ref{prop:whole-ker-near-bound}, which we'll state and prove in Section~\ref{sec:bounds on k}, we can find a neighborhood $\near \subset \domain$ of $\zeta = 0$ with the property that
\begin{equation}\label{near-limit}
|k(a, a')| \le \frac{\sigma}{|\zeta(a)|}
\end{equation}
for all $a, a' \in \near$. By choosing a small enough positive radius $\delta$, we can take $\near$ to be the part of $\domain$ where $|\zeta| < \delta$. Complementarily, let $\far$ be the part of $\domain$ where $\delta \le |\zeta|$.

>>>>

We'll reduce Theorem~\ref{thm:general_volterra} to Lemma~\ref{lem:perturbed_volterra} in Section~\ref{sec:existence and uniqueness}, by rewriting the homogeneous equation we want to solve as an inhomogeneous equation in a more regular space. To show that the inhomogeneous term, $\softpart f_0$, is regular enough, we'll use Proposition~\ref{prop:smoothing} from Section~\ref{sec:image under soft_part} proving that $\softpart$ improves on the regularity of $f_0$ described by Theorem~\ref{thm:proto-growth}.
\begin{rmk}
When $\hardpart$ also satisfies Condition~\eqref{cond:reg-p}, Theorem~\ref{thm:general_volterra} can be restated to give a unique solution in the affine subspace
\[ \zeta^{\tau-1} + \singexpalg{\rho-1}(\domain) \]
when $\rho > \tau$ is low enough, as described in Proposition~\ref{prop:alt-general_volterra} from Section~\ref{sec:existence and uniqueness}.
\end{rmk}
\subsection{Acknowledgements}
We are grateful for the conversations with Kihyun Kim, Angeliki Menegaki, and Sung-Jin Oh that informed the early stages of this work. This paper is a result of the ERC-SyG project, Recursive and Exact New Quantum Theory (ReNewQuantum) which received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme under grant agreement No 810573. It was also made possible by the hospitality of the Institut des Hautes \'{E}tudes Scientifiques during a return visit by AF, funded by the Fondation Math\'{e}matique Jacques Hadamard under the \textit{Junior Scientific Visibility} program.
\section{Function spaces for holomorphic Volterra operators}\label{fn-spaces}
\subsection{Weighted holomorphic $L^{\infty}$ spaces}\label{sec:fn-space-defs}
Throughout this paper, as described in Section~\ref{motivation}, the ``position'' variable $\zeta$ will be the standard coordinate on $\C$. Take a simply connected open set $\domain \subset \C$ that touches but doesn't contain $\zeta = 0$. Let $\cont(\domain)$ be the space of continuous complex-valued functions on $\domain$. Give $\cont(\domain)$ the compact-open topology, recalling that this is the coarsest topology in which the seminorm $f \mapsto \sup_K |f|$ is continuous for every compact subset $K \subset \domain$~\cite[Example~2.6 and Section~4 notes]{fnl-cpx-anal}. The holomorphic functions form a closed subspace $\holo(\domain) \subset \cont(\domain)$~\cite[Proposition~3.14]{fnl-cpx-anal}.

Fixing a real constant $\Lambda$, let's restrict our attention to holomorphic functions on $\domain$ which are bounded by constant multiples of $e^{\Lambda|\zeta|}$. One might describe these functions as being uniformly of exponential type $\Lambda$.\footnote{Recall that a function is of exponential type $\Lambda$ if for every $\varepsilon>0$, there's a constant $A_\varepsilon$---which may depends on $\varepsilon$---such that $|f|\le A_\varepsilon e^{(\Lambda+\varepsilon)|\zeta|}$. We instead require a uniform constant $A$ such that $|f| \le A e^{\Lambda|\zeta|}$.} They form a space $\singexp{0}{\Lambda}(\domain)$, which we'll equip with the norm $\|f\|_{0,\Lambda} = \sup_\domain e^{-\Lambda|\zeta|}\,|f|$. With respect to the seminorm on $\holo(\domain)$ given by a compact set $K \subset \domain$, the inclusion map $\singexp{0}{\Lambda}(\domain) \hookrightarrow \holo(\domain)$ has norm $\sup_K e^{\Lambda |\zeta|}$. That means the inclusion is continuous.
\begin{prop}\label{exp-complete}
The space $\singexp{0}{\Lambda}(\domain)$ is complete.
\end{prop}
\begin{proof}
Take a Cauchy sequence $f_1, f_2, f_3, \ldots \in \singexp{0}{\Lambda}(\Omega)$. The inclusion map $\singexp{0}{\Lambda}(\Omega) \hookrightarrow \holo(\Omega)$ is bounded with respect to each of the seminorms on $\holo(\Omega)$ given by $|f| \mapsto \sup_K |f|$ for compact subsets $K \subset \Omega$, so our sequence is Cauchy in $\holo(\Omega)$ too. Since $\holo(\Omega)$ is complete~\cite[Proposition~3.5]{fnl-cpx-anal},\footnote{That is, a sequence which is Cauchy in each of the seminorms on $\holo(\Omega)$ will always converge in the topology of $\holo(\Omega)$, which is the coarsest topology in which all of the seminorms are continuous.} our sequence converges to a function $f$ there.

The Cauchy property in $\singexp{0}{\Lambda}(\Omega)$ tells us that for any $r > 0$, we can find some $n$ for which $e^{-\Lambda |\zeta|}\,|f_k - f_n| \le r$ whenever $k \ge n$. Since convergence in $\holo(\Omega)$ implies pointwise convergence, we can see as $k$ grows that $e^{-\Lambda |\zeta|}\,|f - f_n| \le r$. This shows that our sequence converges to $f$ in the norm $\|\cdot\|_{0,\Lambda}$. We can also see from this argument that $f$ is in $\singexp{0}{\Lambda}(\domain)$: picking some $r > 0$, we observe that
\begin{align*}
e^{-\Lambda |\zeta|}\,|f| & \le e^{-\Lambda |\zeta|}\,|f - f_n| + e^{-\Lambda |\zeta|}\,|f_n| \\
& \le r + \|f_n\|_{0,\Lambda}
\end{align*}
for the corresponding $n$, showing that $e^{-\Lambda |\zeta|}\,|f|$ is bounded.
\end{proof}

Now, let's relax our norm to allow both exponential growth at infinity and a power-law singularity at $\zeta = 0$. For any $\sigma \in \R$, let $\singexp{\sigma}{\Lambda}(\Omega)$ be the space of holomorphic functions on $\Omega$ which are bounded by constant multiples of $|\zeta|^\sigma e^{\Lambda|\zeta|}$. Give it the norm $\|f\|_{\sigma,\Lambda} = \sup_\Omega |\zeta|^{-\sigma} e^{-\Lambda|\zeta|}\,|f|$. Reprising the arguments from above, we can show that the inclusion $\singexp{\sigma}{\Lambda}(\Omega) \hookrightarrow \holo(\Omega)$ is continuous, and we can generalize Proposition~\ref{exp-complete}:

@@

As we mentioned in Section~\ref{sec:fn-space-defs}, one might describe $\singexp{0}{\Lambda}(\domain)$ as the space of functions on $\domain$ which are uniformly of exponential type $\Lambda$. Taking the union of these spaces over all $\Lambda \in \R$, we get the space $\singexpalg{0}(\domain)$ that contains all functions of exponential type. Having continuous inclusions between the subspaces $\singexp{0}{\Lambda}(\domain)$ as $\Lambda$ increases, we can give $\singexpalg{0}(\domain)$ a meaningful topology: the finest topology that makes all the inclusions $\singexp{0}{\Lambda}(\domain) \hookrightarrow \singexpalg{0}(\domain)$ continuous.\footnote{In category-theoretic language, $\singexpalg{0}(\domain)$ is the limit of the family $\singexp{0}{\Lambda}(\domain)$.} For any $\sigma \in \R$, we define $\singexpalg{\sigma}(\domain)$ similarly.

@@

When $\sigma' > \sigma$, there's a continuous inclusion $\singexpalg{\sigma'}(\domain)\hookrightarrow \singexpalg{\sigma}(\domain)$.

@@

\begin{prop}\label{prop:get-contraction}
Suppose that $\hardpart$ satisfies {\em Conditions \eqref{cond:sing}} and \eqref{cond:diag-basic}, and $\softpart$ satisfies {\em Condition \eqref{cond:eps-lambda}}. Then, for each $\rho > \tau$, we can ensure that $\volterra$ is a contraction of $\singexp{\rho-1}{\Lambda}(\domain)$ by making $\Lambda$ big enough.
\end{prop}
\begin{rmk}
For the argument, we'll use, ``big enough'' always requires $\Lambda > \lambda_\Delta$, and may require $\Lambda \gg \lambda_\Delta$.
\end{rmk}
First, pick some $\sigma \in (\tau, \rho)$.\footnote{If you'd like, you can fix a choice of $\sigma$---for example, taking $\sigma = \frac{1}{2}(\tau + \rho)$.} By Proposition~\ref{prop:whole-ker-near-bound}, which we'll state and prove in Section~\ref{sec:bounds on k}, we can find a neighborhood $\near \subset \domain$ of $\zeta = 0$ with the property that
\begin{equation}\label{near-limit}
|k(a, a')| \le \frac{\sigma}{|\zeta(a)|}
\end{equation}
for all $a, a' \in \near$. By choosing a small enough positive radius $\delta$, we can take $\near$ to be the part of $\domain$ where $|\zeta| < \delta$. Complementarily, let $\far$ be the part of $\domain$ where $\delta \le |\zeta|$.
